---
title: "Lab5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Question 1**
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(ggplot2)
library(dplyr)
library(lubridate)
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
pers <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

fatal <- pers %>% filter(INJ_SEV == "4")
```

**Question 2**
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(ggplot2)
library(dplyr)
library(lubridate)
fatal <- pers %>% filter(INJ_SEV == "4")
dangerousMake <- fatal %>% group_by(MAKE, STATE) %>% 
  summarise(
  personsHit=sum(PER_NO))
```

**Question 3**
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(ggplot2)
library(dplyr)
library(lubridate)
state <- map_data('state')
glc <- readxl::read_xlsx("FRPP_GLC_-_United_StatesNov42021.xlsx")
glc$`State Name`<- tolower(glc$`State Name`)
glc$`State Code`<- as.numeric(glc$`State Code`)
labels <- state %>% group_by(region) %>% 
  summarise(long=mean(long, na.rm=TRUE), lat=mean(lat, na.rm=TRUE))
labels <- left_join(labels, glc, 
                    by=c('region' = 'State Name'))
labels <- left_join(labels, dangerousMake, 
                    by=c('State Code'='STATE'))
labels$MAKE <- factor(labels$MAKE, levels=c(12, 20, 49, 37),
                      labels=c('Ford', 'Chevrolet', 'Toyota', 
                               'Honda'))
state %>% ggplot() + 
  geom_polygon(aes(x=long, y=lat, group=group),
               color='grey', fill='black') + 
  geom_text(data=labels, aes(label=MAKE, x=long, y=lat), 
            color='lightblue', size=3)
```

**Question 4**
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(ggplot2)
library(dplyr)
library(lubridate)
acc2 <- subset(acc, select = 
                 c(STATE, ST_CASE,
                   COUNTY, 
                   CITY, DAY, 
                   MONTH, YEAR, 
                   DAY_WEEK, HOUR, 
                   LATITUDE, LONGITUD))
pers2 <- subset(pers, select = 
                  c(STATE, ST_CASE,
                    COUNTY, DAY,
                    MONTH, HOUR, 
                    MAKE, SEX,
                    INJ_SEV))
joined <- full_join(acc2, pers2, by='ST_CASE')
joinedFatal <- joined %>% filter(INJ_SEV=='4')
```

I decided to drop the columns that seemed unecessary, as RStudio could not create a dataframe that was large enough to contain all of the data.

**Question 5**
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(ggplot2)
library(dplyr)
library(lubridate)
joinedFatal %>% ggplot(aes(x=wday(DAY_WEEK, label = TRUE))) + 
                         geom_bar()
```

The days of the week that have the highest number of deaths related to car accidents are Fridays, Saturdays, and Sundays.

```{r}
joinedFatal %>% ggplot(aes(x=SEX)) + geom_bar()
```

In the dataset, males are coded as 1 and females are coded as 2. According to this graph, men make up much of car accident deaths.

```{r}
joinedFatal %>% ggplot(aes(x=HOUR.x)) + 
  geom_bar()
```

Many car accident deaths take place in the evenings and late at night, there are relatively few in the mornings and there is a steady increase after noon.

**Question 6**
```{r}
ggplot(state, aes(x=long, y=lat)) + 
    geom_polygon(aes(group=group)) + 
    geom_point(aes(x=LONGITUD, y=LATITUDE), data=joinedFatal,
               color='lightblue',
               alpha=0.2, size=0.02) + 
    xlim(-130, -60) + ylim(20, 50) +
    coord_map()
```

Most of the observed deaths were concentrated around urban areas. I noticed that the areas where cities like Chicago, New York City, Las Angeles, Dallas, and Miami are roughy located also have a high number of car accident deaths.
